---
title: "R Notebook"
output: html_notebook
---

## Code for govhack 2016
## Small Business Viability
## Justin Kerney, Ashleigh Latter, Chris Keane

## This code uses Historical New Enterprise Incentive Scheme Business Level Data
## And ABS Labour Force Data

## Creates plots of likelihood of success based on industry, age, location
## plus forecasts of industry size in 2020

## Please excuse the quality of the code. Limited time & sleep.


```{r}

library(tidyverse)
library(ggthemes)
library(scales)
library(stringr)

```

```{r}

inputFile1 <- "/Users/jkerney/Desktop/govhack/neisdatagovhack.csv"
inputFile2 <- "/Users/jkerney/Desktop/govhack/combined.csv"
df <- read.csv(inputFile1)
indData <- read.csv(inputFile2)

```

## Work out success rate for every combination of inputs
```{r}
## Inputs are
## df - the neis data
## user - vector describing user's characteristics

## Outputs are 
## successScore - score out of 10 for likelihood of success

df2 <- df
df2$year <- str_sub(df2$start_date,-2,9)
df2 <- filter(df2, year %in% c(13:17))
user <- c("retail_trade", "SA","0","40_44")

df2$metro <- as.factor(df2$metro)

df2 <- select(df2,
             successful,
             industry_type,
             state,
             metro,
             age_group)

## summarise success rates for each combination of variables
resultsSummary <- group_by(df2, 
                           industry_type,
                           state,
                           metro,
                           age_group) %>%
                  summarise(PercentSuccessful = sum(successful == "Y")/n(), Count = n())

## Work out percentile
resultsSummary <- arrange(resultsSummary, PercentSuccessful)
resultsSummary$rank <- seq(1:nrow(resultsSummary))
resultsSummary$Percentile <- resultsSummary$rank / nrow(resultsSummary) * 100

userResult <- filter(resultsSummary, industry_type == user[1], 
                        state == user[2], 
                        metro == user[3],
                        age_group == user[4])

## Calculate the User's Success Score
successScore <- round(userResult$Percentile/10, 0)
print(successScore)

## Create the dataframe for the first summary plot
Groups <- c("Your Industry,\nState,\nAge Group,\nMetro\\Regional", "Your Industry", "Your Age Group", "Your State", "National")
successPlotDf <- data.frame(group = Groups, successRate = rep(0,5))
successPlotDf$group = factor(successPlotDf$group,levels(successPlotDf$group)[c(4,2,3,5,1)])

## You Result
userResult <- filter(resultsSummary, industry_type == user[1], 
                        state == user[2], 
                        metro == user[3],
                        age_group == user[4])

successPlotDf[1,]$successRate <- userResult$PercentSuccessful

## Your Industry Result
userResult <- filter(df2, industry_type == user[1]) %>%
              summarise(PercentSuccessful = sum(successful == "Y")/n(), Count = n())

successPlotDf[2,]$successRate <- userResult$PercentSuccessful

## Your Age Group Result
userResult <- filter(df2, age_group == user[4]) %>%
              summarise(PercentSuccessful = sum(successful == "Y")/n(), Count = n())

successPlotDf[3,]$successRate <- userResult$PercentSuccessful

## Your State Result
userResult <- filter(df2, state == user[2]) %>%
              summarise(PercentSuccessful = sum(successful == "Y")/n(), Count = n())

successPlotDf[4,]$successRate <- userResult$PercentSuccessful

## National Result
userResult <- summarise(df2, PercentSuccessful = sum(successful == "Y")/n(), Count = n())

successPlotDf[5,]$successRate <- userResult$PercentSuccessful

## Min growth
minGrowth <- min(successPlotDf$successRate) - .05
## Max growth
maxGrowth <- max(successPlotDf$successRate) + .05

## Success Rate
ggplot(data=successPlotDf) +
  geom_bar(mapping=aes(x=group, y=successRate, fill=rainbow(5)), stat="identity") +
  theme_fivethirtyeight() +
  scale_fill_discrete(guide=FALSE) +
  labs(title="Success Rate of NEIS Small Businesses Owners\nSimilar to You, 2013-2017", caption="Data Source: Historical New Enterprise Incentive Scheme Business Level Data") +
  scale_y_continuous(labels=percent) +
  coord_cartesian(ylim=c(0,maxGrowth)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.caption = element_text(size=7, face="italic"))

## Your Industry
successPlotInd <- filter(df2, state== user[2]) %>%
                  group_by(industry_type) %>%
                  summarise(PercentSuccessful = sum(successful == "Y")/n(), Count = n()) %>%
                  arrange(PercentSuccessful)

successPlotInd$niceIndustry <- "test"

## Make industry names nicer
successPlotInd[successPlotInd$industry_type == "electricity_gas_water",]$niceIndustry <- "Electricity, Gas & Water"
successPlotInd[successPlotInd$industry_type == "other",]$niceIndustry <- "Other"
successPlotInd[successPlotInd$industry_type == "transport",]$niceIndustry <- "Transport"
successPlotInd[successPlotInd$industry_type == "finance_and_insurance",]$niceIndustry <- "Finance & Industry"
successPlotInd[successPlotInd$industry_type == "retail_trade",]$niceIndustry <- "Retail/Trade"
successPlotInd[successPlotInd$industry_type == "personal_and_other",]$niceIndustry <- "Personal & Other"
successPlotInd[successPlotInd$industry_type == "health_and_community",]$niceIndustry <- "Health & Community"
successPlotInd[successPlotInd$industry_type == "communication",]$niceIndustry <- "Communication"
successPlotInd[successPlotInd$industry_type == "manufacturing",]$niceIndustry <- "Manufacturing"
successPlotInd[successPlotInd$industry_type == "cultural_and_recreational",]$niceIndustry <- "Cultural & Recreactional"
successPlotInd[successPlotInd$industry_type == "construction",]$niceIndustry <- "Construction"
successPlotInd[successPlotInd$industry_type == "education",]$niceIndustry <- "Education"
successPlotInd[successPlotInd$industry_type == "agriculture_forestry_fishing",]$niceIndustry <- "Agriculture, Forestry & Fishing"
successPlotInd[successPlotInd$industry_type == "accommodation_cafes_restaurants",]$niceIndustry <- "Accomodation, Cafes, Restaurants"
successPlotInd[successPlotInd$industry_type == "wholesale_trade",]$niceIndustry <- "Wholesale Trade"
successPlotInd[successPlotInd$industry_type == "property_and_business",]$niceIndustry <- "Property & Business"

temp <- select(successPlotInd, niceIndustry, PercentSuccessful) %>%
        arrange(PercentSuccessful)

successPlotInd <- within(successPlotInd, 
                   niceIndustry <- factor(niceIndustry, 
                                      levels=temp$niceIndustry))

ggplot(data=successPlotInd) +
  geom_bar(mapping=aes(x=niceIndustry, y=PercentSuccessful, fill=industry_type==user[1]), stat="identity") +
  theme_fivethirtyeight() +
  scale_fill_discrete(guide=FALSE) +
  labs(title="Success Rate of NEIS\nSmall Businesses Owners\nBy Industry For Your State\n2013-2017", caption="Data Source: Historical New Enterprise Incentive Scheme Business Level Data") +
  scale_y_continuous(labels=percent) +
  coord_cartesian(ylim=c(0,maxGrowth)) +
  #theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.caption = element_text(size=7, face="italic")) +
  coord_flip()

```

## Forecast Plots
```{r}

# get data for your industry
indEmpDf <- select(indData, date, one_of(user[1]))
colnames(indEmpDf) <- c("date","numPeople")
indEmpDf$row <- 1:nrow(indEmpDf)
indEmpDf$forecast <- indEmpDf$numPeople
indEmpDf[1:41,]$forecast <- NA
indEmpDf[42:nrow(indEmpDf),]$numPeople <- NA
indEmpDf$month <- rep(c("05","08","11","02"), ceiling(nrow(indEmpDf)/4))[1:53]
indEmpDf$year <- str_sub(indEmpDf$date, -2,6)
indEmpDf$date <- paste("1", indEmpDf$month, indEmpDf$year, sep="-")
indEmpDf$date <- as.Date(indEmpDf$date, format="%d-%m-%y")

ggplot(data=indEmpDf) +
  geom_line(mapping=aes(x=date,y=numPeople), colour="darkblue") +
  geom_line(mapping=aes(x=date,y=forecast), colour="red", linetype="dashed") +
  geom_vline(mapping=aes(xintercept = as.numeric(as.Date("2017-06-15", format="%Y-%m-%d"))), colour="red") +
  theme_fivethirtyeight() +
  labs(title="Your Industry's Forecasted Size in 2020", x="Year",y="Total Employment (thousands)", caption="Data Source: ABS Labour Force Region - SA4 Data") +
  theme(plot.caption = element_text(size=7, face="italic"))

```


